### 2021-04-06 papers/SYRCoSE/future.tex
SYRCoSE: various minor fixes

First, we plan to apply our library to a wider range of problems. 
We are going to develop a mechanized semantics of some classical 
languages used to model concurrency, in particular 
the calculus of communicating systems~(CCS)~\cite{Milner:80} and 
$\pi$-calculus~\cite{Milner:99}.      
We also plan to continue our work on 
expressing various relaxed models of shared memory 
in terms of event structures.

Second, we want to cover other classes of event structures in our library, 
in particular bundle~\cite{Langerak:91}, flow~\cite{Boudol-Castellani:1991}
and stable~\cite{Winskel:82, Winskel:86} event structures.
We plan to use them to develop mechanized \emph{denotational} semantics 
of concurrent languages and relaxed shared memory models.

>>>>

First, we plan to apply our library to a wider range of problems. 
We are going to develop a mechanized semantics of some classical 
languages used to model concurrency, in particular 
the calculus of communicating systems~(CCS)~\cite{Milner:80} and 
$\pi$-calculus~\cite{Milner:99}.      
We also plan to continue our work on 
expressing various relaxed models 
of shared memory~\cite{Sewell-al:CACM10, Lahav-al:POPL16, Lahav-al:PLDI17, Podkopaev-al:POPL19}
in terms of event structures.

Second, we want to cover other classes of event structures in our library, 
in particular bundle~\cite{Langerak:91}, flow~\cite{Boudol-Castellani:1991}
and stable~\cite{Winskel:82, Winskel:86} event structures.
We plan to use them to develop mechanized \emph{denotational} semantics 
of concurrent languages and relaxed shared memory models~\cite{Dodds-al:ESOP18}.

### 2021-04-06 papers/SYRCoSE/regmachine-eventstruct.tex
SYRCoSE: various minor fixes

In this section we present alternative operational semantics
which constructs a prime event structure encoding a 
behavior of the register machine.

@@

Following the theory of axiomatic weak memory models~\cite{\todo{}},
we define the causality relation of the register machine's event structure
as a transitive closure of two relations~---~\emph{program order} 
and \emph{reads-from}, denoted as $\lPO$ and $\lRF$ correspondingly.

@@

In order to construct $\lPO$ and $\lRF$ incrementally
we represent them via their inverse covering functions.

@@

  \lessdot_{\lPO} &\defeq& {\frel{f_{\lPO}}}^{-1}  & \lessdot_{\lRF} &\defeq& {\frel{f_{\lRF}}}^{-1} \\
  \lPO           &\defeq& \lessdot_{\lPO}^+      & \lRF           &\defeq& \lessdot_{\lRF}       \\

@@

The primitive conflict relation is generated by the $f_{\lPO}$ function.
The two events are considered to be in primitive conflict if they are
not equal and have a commond $\lPO$ predecessor. 
For this definition to work properly, we also need 
to assume that in the event structure each thread has 
a special initial event labeled by a distinguished 
\emph{thread start} label $\tslab{}$.

$$ e_1 \primcfRel e_2 \iff e_1 \neq e_2 \wedge f_{\lPO}(e_1) = f_{\lPO}(e2) $$

Finally, we need a way to reconcile the event structure
with the states of machine's threads. 
To do so we use a function $\thrdmap : \eventSet \fun \ThrdState$
which maps event to a thread state obtained as a result of
execution of event's side-effect.

Let us consider an example. Given the program below, 
our semantics builts the event structure as 
shown on \cref{fig:regmachine-eventstruct-example}.

@@

The rules of operational semantics constructing 
the event structure are presented on~\cref{fig:eventstruct-sem}.
The first auxiliary rule \ESAddEventRule adds a new event, sets its 
label, $\lPO$ and $\lRF$ predecessors. 
The \ESIdleRule handles the case when a thread of 
the register machine performs an internal step 
without any side effect. 
It chooses an event $e$ together with 
the thread state $s$ corresponding to it
and performs one step reduction to new state $s'$.
It then updates the mapping of events to thread states.   
The last two rules \ESStoreRule and \ESLoadRule
correspond to store and load perfromed by some thread.  
Similarly to \ESIdleRule, an event $e_{\lPO}$ is selected
and one reduction is performed from the corresponding thread state $s$.
Unlike the \ESIdleRule case, however, a new event $e$ is also generated.
In case of \ESLoadRule additionally an event $e_{\lRF}$ is selected,
such that it has a write label matching the read label of the new event.

>>>>

In this section we present operational semantics
which constructs a prime event structure encoding a 
behavior of the register machine.

@@

We will encode the event structure as a tuple 
$\tup{\eventSeq, \lLAB, \lfPO, \lfRF}$, 
below we explain the meaning of each component, 
and how they together form a prime event structure.

The first component $\eventSeq$ is a sequence of events 
$e_1 \identOrdDual \dots \identOrdDual e_n$
in the order reverse to the the one 
in which events are added to the structure. 
The second component is labeling function $\lLAB : \eventSet \fun \Lab$, 
assigning a label to each event.

Next, following the theory of axiomatic weak memory models~\cite{\todo{}},
we define the causality relation of the register machine's event structure
as a transitive closure of two relations~---~\emph{program order} 
and \emph{reads-from}, denoted as $\lPO$ and $\lRF$ correspondingly.

@@

In order to construct $\lPO$ and $\lRF$ incrementally
we represent them via their inverse covering functions $\lfPO$ and $\lfRF$.

@@

  \lessdot_{\lPO} &\defeq& {\frel{\lfPO}}^{-1}  & \lessdot_{\lRF} &\defeq& {\frel{\lfRF}}^{-1} \\
  \lPO           &\defeq& \lessdot_{\lPO}^+      & \lRF           &\defeq& \lessdot_{\lRF}       \\

@@

The primitive conflict relation is generated by the $\lfPO$ function.
The two events are considered to be in primitive conflict if they are
not equal and have a commond $\lPO$ predecessor. 
For this definition to work properly, we also need 
to assume that in each thread has a special initial event 
labeled by a distinguished \emph{thread start} label $\tslab{}$.

$$ e_1 \primcfRel e_2 \iff e_1 \neq e_2 \wedge f_{\lfPO}(e_1) = f_{\lfPO}(e2) $$

Finally, we need a way to reconcile the event structure
with the states of machine's threads. 
To do so we use a function ${\thrdmap : \eventSet \fun \ThrdState}$
which maps event to a thread state obtained as a result of
execution of event's side-effect.

Let us consider an example. Given the program below, 
our semantics builds the event structure as 
shown on \cref{fig:regmachine-eventstruct-example}.

@@

The rules of operational semantics constructing 
the event structure are presented on~\cref{fig:eventstruct-sem}.
The first auxiliary rule \ESAddEventRule adds a new event, sets its 
label, $\lPO$ and $\lRF$ predecessors. 
The \ESIdleRule handles the case when a thread of 
the register machine performs an internal step 
without any side effect. 
It chooses an event $e$ together with 
the thread state $s$ corresponding to it
and performs one step reduction to a new state $s'$.
It then updates the mapping of events to thread states.   
The last two rules \ESStoreRule and \ESLoadRule
correspond to store and load perfromed by some thread.  
Similarly to \ESIdleRule, an event $e_{\lPO}$ is selected
and one reduction is performed from the corresponding thread state $s$.
Unlike the \ESIdleRule case, however, a new event $e$ is also generated.
In case of \ESLoadRule additionally an event $e_{\lRF}$ is selected,
such that it has a write label matching the read label of the new event.

### 2021-04-06 papers/SYRCoSE/regmachine.tex
SYRCoSE: various minor fixes

  I &\in& \Inst     &::=&                                 & \text{instruction}             \\
              & &   & | & \assignInst{r}{v}               & \text{assign to register}      \\
              & &   & | & \opInst{\otimes}{r_1}{r_2}{r_3}  & \text{apply binary operation}  \\
              & &   & | & \jumpInst{r}{\ip{i}}            & \text{conditional jump}        \\
              & &   & | & \readInst{r}{x}                 & \text{read from memory}        \\
              & &   & | & \writeInst{x}{v}                & \text{write to memory}         \\

  r      &\in& \Reg &   &                                & \text{thread-local register}   \\ 
  x      &\in& \Loc &   &                                & \text{shared memory location}  \\ 
  v      &\in& \Z   &   &                                & \text{value}                   \\
  \ip{i} &\in& \N   &   &                                & \text{instruction label}       \\

@@

We first present the semantics of a single thread program.
In this semantics memory access instructions 
does not actually operate on shared memory,
but rather produce a \emph{label}
denoting the side-effect of the operation 
(see \cref{fig:label-syntax}).
This encoding allows us to decouple 
the semantics of the register machine 
from a \emph{memory model}~\cite{\todo{}}.

@@

The semantics is given in a form of 
\emph{labeled transition system}~\cite{\todo{}}.
State of the system consists of instruction pointer
and a map from registers to their values, 
as shown in \cref{fig:regmachine-thrdstate}. 
The rules of the semantics are standard 
(see \cref{fig:regmachine-thrdsem}).

>>>>

  I &\in& \Inst     &::=&                                 & \text{instruction}             \\
              & &   & | & \assignInst{r}{v}               & \text{assign to register}      \\
              & &   & | & \opInst{\otimes}{r_1}{r_2}{r_3}  & \text{apply binary operation}  \\
              & &   & | & \jumpInst{r}{\ip{i}}            & \text{conditional jump}        \\
              & &   & | & \exitInst                       & \text{exit}                    \\
              & &   & | & \readInst{r}{x}                 & \text{read from memory}        \\
              & &   & | & \writeInst{x}{v}                & \text{write to memory}         \\

  r      &\in& \Reg     &   &                    & \text{thread-local register}   \\ 
  x      &\in& \Loc     &   &                    & \text{shared memory location}  \\ 
  v      &\in& \Z       &   &                    & \text{value}                   \\
  \otimes&\in& \BinOp   &   &                    & \text{binary operation}        \\
  \ip{i} &\in& \N       &   &                    & \text{instruction label}       \\

@@

We first present the semantics of a single thread program.
In this semantics memory access instructions 
does not actually operate on shared memory,
but rather produce a \emph{label}
denoting the side-effect of the operation 
(see \cref{fig:label-syntax}).
This encoding allows us to decouple 
the semantics of the register machine 
from a \emph{memory model}.

@@

The semantics is given in a form of 
\emph{labeled transition system}~%
\footnote{As we have mentioned, in our \coq development we actually use 
the monadic encoding of the operational semantics. 
The labeled transition system can be deduced from this encoding.}
State of the system consists of instruction pointer
and a map from registers to their values, 
as shown in \cref{fig:regmachine-thrdstate}. 
The rules of the semantics are standard 
(see \cref{fig:regmachine-thrdsem}).
